---
title: "Gloeo Data Exploration"
author: "J. Brentrup"
date: "1/21/2022"
output: 
  html_document: 
    fig_caption: yes
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages
```{r}
pacman::p_load(tidyverse, lubridate, openair, knitr)
```


# Read in gloeo data 2005-2016
```{r}
gloeo <- read_csv("00_Data_files/EDI_data_clones/weekly_surface_gloeo_4sites_2005_2016_v10April2020.csv",
      col_types = list(
      date = col_date(format = ""),
      site = col_character(),
      year = col_double(),
      dayofyr = col_double(),
      n_sample = col_double(),
      coloniesperL = col_double(),
      filbundperL = col_double(),
      totalperL = col_double()
    )
  )

```

# Data wrangling
```{r}
# Create tibble with rows for missing observations in 2009-2016 
date <- c(rep(c("2009-06-11","2009-09-28","2015-06-25","2016-06-09","2016-08-10"),times = 1,each = 4),"2014-07-17","2015-06-11","2015-07-23","2015-08-20","2016-08-04")

site <- c(rep(c("NorthSunapeeHarbor","SouthOfTheFells","HerrickCoveSouth","Newbury"),times = 5),"Newbury","SouthOfTheFells","HerrickCoveSouth","SouthOfTheFells","NorthSunapeeHarbor")

odd_obs <- as_tibble(cbind(site,date)) %>%
  mutate(date = as.Date(date)) %>%
  mutate(year = year(date), dayofyr = yday(date)) %>%
  mutate(n_sample = NA, coloniesperL = NA, filbundperL = NA, totalperL = NA)


# eliminate extra sampling day from original data file
gloeo1 <- gloeo %>%
  filter(!date %in% as.Date(c("2009-07-11")))#get rid of extra sampling days

# combine missing and averaged duplicate observations with original data file
# assign "sampling season weeks" numbering 1-20 each year to dates
gloeo2 <- bind_rows(gloeo1, odd_obs) %>%
  mutate(week = week(date)) %>%
  filter(year %in% 2009:2016 & (week %in% 21:40 | dayofyr == 283))%>% #limit to study period
  arrange(date, site) %>%
  select(date, site, coloniesperL, filbundperL, totalperL, year, dayofyr) %>% #keep day of year too
  mutate(season_week = rep(c(1:20),times = 8, each = 4))
  
# Add in month and ln of gloeo
gloeo_final <- gloeo2 %>%
  mutate(month = month(date)) %>%
  select(date,year,month,dayofyr,season_week,site,coloniesperL,filbundperL,totalperL) %>% # rearrange columns to have date items together
  mutate(ln_totalperL = log(totalperL + (1/141.4)))# add ln of gloeo, Total per L is Volume of 2, ~1 m net tows = 141.4 L
# convert to natural log using  log + detection limit = 1/141.4

```

# Plot Gloeo data over time - y-scale varies
Y-scale set to free for each year
```{r}
names(gloeo_final)

gloeo_final <- gloeo_final %>% 
  mutate(month2 = month(date, label = T, abbr = T)) %>% 
  mutate(month_day = mday(date)) %>% 
  mutate(month_day2 = paste0(month_day, "-", month2))

str(gloeo_final)

ggplot(data = gloeo_final, aes(x = month2, y = totalperL, color = site))+
  geom_point(size = 4)+
  geom_line(size = 2)+
  scale_color_manual(values = c("#33a02c", "#b2df8a","#a6cee3", "#1f78b4"))+
  facet_wrap(~year, scales = "free_y")+
  theme_bw()+ #base_family = "Helvetica"
        theme(axis.title.x = element_text(size=22),
              axis.title.y = element_text(size=22),
              #legend.title = element_text(size=22),
              axis.text.y = element_text(size=18, color="black",margin = margin(l=10,r=5, unit = "pt")),
              axis.text.x = element_text(size=18, color="black",margin = margin(b=10,t=5, unit = "pt")),
              #axis.ticks.x = 
              strip.text = element_text(size = 20, hjust = 0.03),
              strip.placement = "outside",
              legend.title = element_text(size = 18),
              #legend.position = "top", #c(0.77,0.83)  
              legend.text = element_text(size=18,color="black", margin = margin(r=7,l=5, unit = "pt")))+
              #legend.spacing.x = unit(0.25,"cm"))+
        #legend.box.background = element_blank())+
        #legend.position = "top")+
        theme(#axis.line = element_line(color="black"),
                plot.background = element_blank(),
                panel.border = element_rect(color = "black", size =1),
                panel.grid.major = element_blank(),
                panel.grid.minor = element_blank(),
                strip.background = element_blank())+
        theme(axis.ticks.length = unit(.2, "cm"))+
        labs(y = "G. echinulata Density (colonies/L)", x=NULL, color = "Site")

  ggsave("gloeo_allsites.pdf", width = 15, height = 8.5, useDingbats=FALSE, dpi = 600)

  ggsave("gloeo_month.pdf", width = 15, height = 8.5, useDingbats=FALSE, dpi = 600)


```

# Plot Gloeo for Map - sum of gloeo @ each site
```{r}
# drop 2014 & 2015 and Fichter
  gloeo3 <- gloeo_final %>% 
  filter(year!= 2015 & year!=2016) %>% 
  filter(site!="SouthOfTheFells") %>% 
    filter(site == "HerrickCoveSouth")

gloeo_sum <- gloeo3 %>% 
  group_by(site, year) %>% 
  summarize(gloeo_sum = sum(totalperL, na.rm = T))

# all sites together
ggplot(data = gloeo_sum, aes(x = year, y = gloeo_sum))+
  geom_bar(stat = "identity", fill = "darkgreen") +
 # scale_color_manual(values = c("#33a02c", "#b2df8a","#a6cee3", "#1f78b4"))+
  facet_wrap(~site)+ #, scales = "free_y"
  theme_bw()+ #base_family = "Helvetica"
        theme(axis.title.x = element_text(size=22),
              axis.title.y = element_text(size=22),
              #legend.title = element_text(size=22),
              axis.text.y = element_text(size=18, color="black",margin = margin(l=10,r=5, unit = "pt")),
              axis.text.x = element_text(size=18, color="black",margin = margin(b=10,t=5, unit = "pt")),
              #axis.ticks.x = 
              strip.text = element_text(size = 20, hjust = 0.03),
              strip.placement = "outside",
              legend.title = element_text(size = 18),
              #legend.position = "top", #c(0.77,0.83)  
              legend.text = element_text(size=18,color="black", margin = margin(r=7,l=5, unit = "pt")))+
              #legend.spacing.x = unit(0.25,"cm"))+
        #legend.box.background = element_blank())+
        #legend.position = "top")+
        theme(#axis.line = element_line(color="black"),
                plot.background = element_blank(),
                panel.border = element_rect(color = "black", size =1),
                panel.grid.major = element_blank(),
                panel.grid.minor = element_blank(),
                strip.background = element_blank())+
        theme(axis.ticks.length = unit(.2, "cm"))+
        labs(y = "Annual Sum of G. echinulata Density (colonies/L)", x=NULL, color = "Site")

# By site bar plot
gloeo_sum_site <- gloeo_sum %>% 
  filter(site == "NorthSunapeeHarbor")

gloeo_label <- c(expression(paste(italic("G. echinulata"), " Density ", "(colonies L" ^-1, ")"))) 


ggplot(data = gloeo_sum_site, aes(x = year, y = gloeo_sum))+
  geom_bar(stat = "identity", fill = "mediumaquamarine") + #mediumaquamarine
 # scale_color_manual(values = c("#33a02c", "#b2df8a","#a6cee3", "#1f78b4"))+
#  facet_wrap(~site)+ #, scales = "free_y"
  scale_x_continuous(breaks = seq(2009,2014, by = 1))+
  #scale_y_continuous(limits = c(0,105), breaks = seq(0,100,by = 25))+
  scale_y_continuous(limits = c(0,20), breaks = seq(0,20,by = 5))+
  theme_bw()+ #base_family = "Helvetica"
        theme(axis.title.x = element_text(size=22),
              axis.title.y = element_text(size=14),
              title = element_text(size=16),
              axis.text.y = element_text(size=12, color="black",margin = margin(l=10,r=5, unit = "pt")),
              axis.text.x = element_text(size=12, color="black",margin = margin(b=10,t=5, unit = "pt")),
              legend.title = element_text(size = 18),
              #legend.position = "top", #c(0.77,0.83)  
              legend.text = element_text(size=18,color="black", margin = margin(r=7,l=5, unit = "pt")))+
        theme(#axis.line = element_line(color="black"),
                plot.background = element_blank(),
                #panel.border = element_blank(), # L border
                #axis.line = element_line(color = "black"),
                panel.border = element_rect(color = "black", size =1), # full box border
                panel.grid.major = element_blank(),
                panel.grid.minor = element_blank(),
                strip.background = element_blank())+
        theme(axis.ticks.length = unit(.2, "cm"))+
        labs(y = gloeo_label, x=NULL) #Herrick Cove South Newbury, NorthSunapeeHarbor , title = "Herrick Cove South" "G. echinulata \n Density (colonies/L)"

  ggsave("~/Documents/Gloeo Bayesian Modeling/Site Model/MS/Figures/gloeo_sum_HC.pdf", width = 4, height = 3.5, useDingbats=FALSE, dpi = 600)
  ggsave("~/Documents/Gloeo Bayesian Modeling/Site Model/MS/Figures/gloeo_sum_NB.pdf", width = 4, height = 3.5, useDingbats=FALSE, dpi = 600)
  ggsave("~/Documents/Gloeo Bayesian Modeling/Site Model/MS/Figures/gloeo_sum_NSH.pdf", width = 4, height = 3.5, useDingbats=FALSE, dpi = 600)
  ggsave("~/Documents/Gloeo Bayesian Modeling/Site Model/MS/Figures/gloeo_sum_NSH_scale.pdf", width = 4, height = 3.5, useDingbats=FALSE, dpi = 600)

# Histogram
  
  gloeo3 <- gloeo_final %>% 
  filter(year!= 2015 & year!=2016) %>% 
  filter(site!="SouthOfTheFells") %>% 
    filter(site == "HerrickCoveSouth")

gloeo_label_log <- c(expression(paste("log(", italic("G. echinulata"), " Density)"))) 

  ggplot(data = gloeo3, aes(x = ln_totalperL))+
  geom_histogram(binwidth = 0.5, fill = "mediumaquamarine", color = "black") + #mediumaquamarine
 # scale_color_manual(values = c("#33a02c", "#b2df8a","#a6cee3", "#1f78b4"))+
#  facet_wrap(~site)+ #, scales = "free_y"
  scale_x_continuous(breaks = seq(-5,5, by = 2.5))+
  scale_y_continuous(limits = c(0,25), breaks = seq(0,25,by = 5))+
  theme_bw()+ #base_family = "Helvetica"
        theme(axis.title.x = element_text(size=22),
              axis.title.y = element_text(size=14),
              title = element_text(size=16),
              axis.text.y = element_text(size=12, color="black",margin = margin(l=10,r=5, unit = "pt")),
              axis.text.x = element_text(size=12, color="black",margin = margin(b=10,t=5, unit = "pt")),
              legend.title = element_text(size = 18),
              #legend.position = "top", #c(0.77,0.83)  
              legend.text = element_text(size=18,color="black", margin = margin(r=7,l=5, unit = "pt")))+
        theme(#axis.line = element_line(color="black"),
                plot.background = element_blank(),
                #panel.border = element_blank(), # L border
                #axis.line = element_line(color = "black"),
                panel.border = element_rect(color = "black", size =1), # full box border
                panel.grid.major = element_blank(),
                panel.grid.minor = element_blank(),
                strip.background = element_blank())+
        theme(axis.ticks.length = unit(.2, "cm"))+
        labs(y = gloeo_label_log, x=NULL) #Herrick Cove South Newbury, NorthSunapeeHarbor , title = "Herrick Cove South" "G. echinulata \n Density (colonies/L)"


  ggsave("~/Documents/Gloeo Bayesian Modeling/Site Model/MS/Figures/gloeo_hist_HC.pdf", width = 4, height = 3.5, useDingbats=FALSE, dpi = 600)
  ggsave("~/Documents/Gloeo Bayesian Modeling/Site Model/MS/Figures/gloeo_hist_NB.pdf", width = 4, height = 3.5, useDingbats=FALSE, dpi = 600)
  ggsave("~/Documents/Gloeo Bayesian Modeling/Site Model/MS/Figures/gloeo_hist_NSH.pdf", width = 4, height = 3.5, useDingbats=FALSE, dpi = 600)

```

# Plot Gloeo data over time - y-scale fixed
Y-scale set to max of 45 to exclude 80 for Midge
```{r}
ggplot(data = gloeo_final, aes(x = dayofyr, y = totalperL, color = site))+
  geom_point()+
  geom_line()+
  scale_y_continuous(limits = c(0,45), breaks = seq(0,45,by = 5))+
  facet_wrap(~year) +
  labs(title = "Y-Scale Fixed Gloeo 2009-2016")

```

# Plot LN Gloeo data over time - y-scale fixed
```{r}
ggplot(data = gloeo_final, aes(x = dayofyr, y = ln_totalperL, color = site))+
  geom_point()+
  geom_line()+
 # scale_y_continuous(limits = c(0,45), breaks = seq(0,45,by = 5))+
  facet_wrap(~year)+
  labs(title = "LN Gloeo 2009-2016")
```

# Summary of gloeo data for calibration period 2009-2014 ####
```{r}
gloeo_cal <- gloeo_final %>% 
  filter(year < 2015) %>% 
  group_by(site, year) %>% 
  summarize_at(vars(totalperL), list(~ mean(., na.rm = TRUE), ~ median(., na.rm = TRUE),~ min(., na.rm = TRUE),~ max(., na.rm = TRUE), ~ sd(., na.rm = TRUE), ~ IQR(., na.rm = TRUE), ~n()))

gloeo_cal

kable(gloeo_cal, digits = 2, caption = "Gloeo Data Summary - All Sites & Years")

gloeo_cal_v2 <- gloeo_final %>% 
  filter(year < 2015) %>% 
  group_by(site) %>% 
  summarize_at(vars(totalperL), list(~ mean(., na.rm = TRUE), ~ median(., na.rm = TRUE),~ min(., na.rm = TRUE),~ max(., na.rm = TRUE), ~ sd(., na.rm = TRUE), ~ IQR(., na.rm = TRUE), ~n()))

gloeo_cal_v2

kable(gloeo_cal_v2, digits = 2, caption = "Gloeo Data Summary - All Sites")

gloeo_cal_ln <- gloeo_final %>% 
  filter(year < 2015) %>% 
  group_by(site) %>% 
  summarize_at(vars(ln_totalperL), list(~ mean(., na.rm = TRUE), ~ median(., na.rm = TRUE),~ min(., na.rm = TRUE),~ max(., na.rm = TRUE), ~ sd(., na.rm = TRUE), ~ IQR(., na.rm = TRUE), ~n()))

gloeo_cal_ln

kable(gloeo_cal_ln, digits = 2, caption = "LN Gloeo Data Summary - All Sites")

```

## Pre 2009 data for observation error prior
- Filter data for pre 2009 to look at mean + variability for each site
- Midge data 2005-2008
- Newbury data 2005, 2007-2008
- Fichter/Coffin 2007-2008

```{r}
unique(gloeo$site)

gloeo_pre_2009 <- gloeo %>% 
  filter(year < 2009) %>% 
  group_by(site, year) %>% 
  summarize_at(vars(totalperL), list(~ mean(., na.rm = TRUE), ~ median(., na.rm = TRUE),~ max(., na.rm = TRUE), ~ sd(., na.rm = TRUE), ~ IQR(., na.rm = TRUE), ~n()))

gloeo_pre_2009

kable(gloeo_pre_2009, digits = 2, caption = "Pre-2009 Gloeo Data Summary - All Sites")

# Natural log gloeo data
ln_gloeo_pre_2009 <- gloeo %>% 
   mutate(ln_totalperL = log(totalperL + (1/141.4))) %>% # add ln of gloeo, Total per L is Volume of 2, ~1 m net tows = 141.4 L convert to natural log using  log + detection limit = 1/141.4
  filter(year < 2009) %>% 
  group_by(site, year) %>% 
  summarize_at(vars(ln_totalperL), list(~ mean(., na.rm = TRUE), ~ median(., na.rm = TRUE),~ max(., na.rm = TRUE), ~ sd(., na.rm = TRUE), ~ IQR(., na.rm = TRUE), ~n()))

ln_gloeo_pre_2009

kable(ln_gloeo_pre_2009, digits = 2, caption = "Pre-2009 LN Gloeo Data Summary - All Sites")

# Filter for early season data only - 2008 earliest 1st week of June
ln_gloeo_2008_June <- gloeo %>% 
   mutate(ln_totalperL = log(totalperL + (1/141.4))) %>% # add ln of gloeo, Total per L is Volume of 2, ~1 m net tows = 141.4 L convert to natural log using  log + detection limit = 1/141.4
  filter(year == 2008) %>% 
  mutate(month = month(date)) %>% 
  filter(month == 6) %>% 
  filter(date < "2008-06-13")

summary(ln_gloeo_2008_June) # Median :0.01400  ln: -3.860

ggplot(ln_gloeo_2008_June, aes(x = ln_totalperL))+
  geom_density()

ln_gloeo_2008_June_summary <- gloeo %>% 
   mutate(ln_totalperL = log(totalperL + (1/141.4))) %>% # add ln of gloeo, Total per L is Volume of 2, ~1 m net tows = 141.4 L convert to natural log using  log + detection limit = 1/141.4
  filter(year == 2008) %>% 
  mutate(month = month(date)) %>% 
  filter(month == 6) %>% 
  filter(date < "2008-06-13") %>% 
   summarize_at(vars(ln_totalperL), list(~ mean(., na.rm = TRUE), ~ median(., na.rm = TRUE),~ min(., na.rm = TRUE),~ max(., na.rm = TRUE), ~ sd(., na.rm = TRUE), ~ IQR(., na.rm = TRUE), ~n()))

kable(ln_gloeo_2008_June_summary, digits = 2, caption = "June 2008 LN Gloeo Data Summary - All Sites")

```

# Histograms
```{r}
# Gloeo colonies - no log
ggplot(data = gloeo_final, aes(x = totalperL)) +
  geom_histogram(color = "white")+
  facet_wrap(~site, scales = "free_x")

# Gloeo colonies - log
ggplot(data = gloeo_final, aes(x = ln_totalperL)) +
  geom_histogram(binwidth = 1, color = "white")+
  facet_wrap(~site)

```

# Viz by site and years overlayed
```{r}
gloeo_viz <- gloeo_final %>% 
  mutate(Year = as.factor(year(date)))

ggplot(gloeo_viz, aes(x = season_week, y = totalperL, group = Year, color = Year))+
  geom_line(size = 1.5)+
  geom_point(size = 2) +
 # scale_y_continuous(limits = c(0,45), breaks = seq(0,45,by = 5))+
  scale_color_brewer(palette = "BrBG")+
  theme_classic()+
  labs(x = "Week of sampling season", y = "Gloeo (colonies per L)", title = "All Sites- Y Scale Varies")+
  theme(axis.title = element_text(size = 16),axis.text = element_text(size = 14),legend.title = element_text(size = 16),legend.text = element_text(size = 14))+
  facet_wrap(~site, scales = "free_y")

```


# Water Temp data from Onset Loggers @ each site
## Load onset water temp data into R 
```{r}
wtr <- read_csv("./00_Data_files/EDI_data_clones/temp_2006-2018_QAQC_vert_09May2020.csv",
                col_types = list(year = col_double(),
                                 dayofyr = col_double(),
                                 time = col_time(),
                                 datetime = col_datetime(),
                                 site = col_character(),
                                 temp_degC = col_double()))
```

## Data wrangling - all sites

- limit logger data to sampling years & sampling weeks and remove 30 min data in 2009 so consistent with other years

```{r}
wtr1 <- wtr %>%
  mutate(date = date(datetime)) %>%
  filter(year %in% 2009:2016) %>%
  mutate(min = minute(time)) %>%
  filter(min!=30) %>%
  select(-min) %>%
  mutate(week = week(datetime)) %>%
  filter(week %in% 21:40 | dayofyr == 283) %>% 
  select(datetime, date, year, dayofyr, time, week, site, temp_degC)
```

## Plot water temp data over time - y-scale varies
Y-scale set to free for each year
```{r}
names(wtr1)
# Remember hourly data - so plots look a little funky with dayofyr on x-axis but can do facet
ggplot(data = wtr1, aes(x = dayofyr, y = temp_degC, color = site))+
  #geom_point()+
  geom_line()+
  facet_wrap(~year, scales = "free_y")
```

## Plot water temp data over time - y-scale fixed
Y-scale set to min of 10 and max of 30 max = 28.62 @ Newbury 2013 July 19
```{r}
summary(wtr1)

# Remember hourly data - so plots look a little funky with dayofyr on x-axis
ggplot(data = wtr1, aes(x = dayofyr, y = temp_degC, color = site))+
  #geom_point()+
  geom_line()+
  scale_y_continuous(limits = c(10,30), breaks = seq(10,30,by = 5))+
  facet_wrap(~year)
```

### Remove outlier values
```{r}
# High values @ North Sunapee Harbor 2013 pre DOY 160 start after 2013-05-30 11:00:00
# 2014 - ~DOY 260 spike @ Fichter last value 2014-09-17 09:00:00, Coffin last value same
sum(is.na(wtr1)) #1348

# Plots zoomed in by year and site
wtr_site <- wtr1 %>% 
  filter(year == 2013) %>% 
 # filter(dayofyr < 153) %>% 
  filter(site == "NorthSunapeeHarbor")

# Hourly dataset - plot with datetime on x-axis
ggplot(data = wtr_site, aes(x = datetime, y = temp_degC, color = site))+
  geom_point() +
  geom_line()

# Jagged warm up @ Newbury at beginning of season for many years

# Clean data remove high values @ Fichter and Coffin in 2013 & 2014
wtr2 <- wtr1 %>% 
  mutate(temp_degC = ifelse(site == "NorthSunapeeHarbor" & datetime >= "2013-05-23 11:00:00" & datetime < "2013-05-30 7:00:00", NA, temp_degC)) %>% 
  mutate(temp_degC = ifelse(site == "NorthSunapeeHarbor" & datetime >= "2014-09-17 05:00:00" & datetime < "2014-09-30 7:00:00", NA, temp_degC)) %>% 
  mutate(temp_degC = ifelse(site == "SouthOfTheFells" & datetime >= "2014-09-17 05:00:00" & datetime < "2014-09-30 7:00:00", NA, temp_degC))

#write.csv(wtr2, "00_Data_files/EDI_data_clones/wtrtemp_clean_2009-2016_allsites.csv", row.names = F)
```

### Difference in hourly water temp between Midge & Newbury
```{r}
# Calculate difference in water temp between Midge - Newbury for hourly water temp data
wtr_diff <- wtr2 %>% 
  filter(site == "HerrickCoveSouth"|site == "Newbury") %>% 
  pivot_wider(names_from = site, values_from = temp_degC) %>% 
  mutate(wtr_diff = HerrickCoveSouth - Newbury)
 
# Plot water temp diff over time 
ggplot(data = wtr_diff, aes(x = dayofyr, y = wtr_diff))+
  #geom_point(color = "blue") +
  geom_line() +
  scale_y_continuous(limits = c(-5.3,9), breaks = seq(-5,9,by = 2))+
  geom_hline(yintercept = 0, linetype = 2, color = "red")+
  facet_wrap(~year)

summary(wtr_diff) # max = 8.85 deg difference - wow
  
```

### Temp visualization
```{r}
wtr2_viz <- wtr2 %>% 
  group_by(site, date) %>% 
  summarize(temp_degC_mean = mean(temp_degC, na.rm = T)) %>% 
  mutate(doy = yday(date), Year = as.factor(year(date)))

# Facet by sites
ggplot(wtr2_viz, aes(x = doy, y = temp_degC_mean, group = Year, color = Year))+
  geom_line(size = 1.3)+
  scale_color_brewer(palette = "PRGn")+
  theme_classic()+
  xlab("DOY")+
  ylab("Water temp. (°C)")+
  theme(axis.title = element_text(size = 16),axis.text = element_text(size = 14),legend.title = element_text(size = 16),legend.text = element_text(size = 14),strip.text = element_text(size = 14))+
  facet_wrap(~site)

# Facet by year
ggplot(wtr2_viz, aes(x = doy, y = temp_degC_mean, group = site, color = site))+
  geom_line(size = 1.3)+
  scale_color_brewer(palette = "Spectral")+
  theme_classic()+
  xlab("DOY")+
  ylab("Water temp. (°C)")+
  theme(axis.title = element_text(size = 16),axis.text = element_text(size = 14),legend.title = element_text(size = 16),legend.text = element_text(size = 14),strip.text = element_text(size = 14))+
  facet_wrap(~Year)

```

### Difference in daily water temp between Midge & Newbury

```{r}
# Calculate difference in water temp between Midge - Newbury for hourly water temp data
wtr_diff_daily <- wtr2_viz %>% 
  filter(site == "HerrickCoveSouth"|site == "Newbury") %>% 
  pivot_wider(names_from = site, values_from = temp_degC_mean) %>% 
  mutate(wtr_diff = HerrickCoveSouth - Newbury) %>% 
  mutate(year = year(date))
 
# Plot water temp diff over time 
ggplot(data = wtr_diff_daily, aes(x = doy, y = wtr_diff))+
  geom_point(color = "blue") +
  geom_line() +
 # scale_y_continuous(limits = c(-5.3,9), breaks = seq(-5,9,by = 2))+
  geom_hline(yintercept = 0, linetype = 2, color = "red")+
  facet_wrap(~year)

summary(wtr_diff_daily) # max = 7 deg difference - wow

```

# Temp effects from model parameters
```{r}
temp_effects <- tibble(site = c("Herrick Cove", "Newbury", "North Sunapee Harbor"), wtr_temp_mean = c(0.3, 0.72, 0.4), wtr_temp_sd = c(0.27, 0.26, 0.21),gdd_mean = c(0.19, 0.03, 0.05), gdd_sd = c(0.16, 0.17, 0.12), gdd_quad_mean = c(-0.48, -0.08, -0.21), gdd_quad_sd = c(0.32, 0.28, 0.24))

#Min Water Temp 
ggplot(temp_effects, aes(x = wtr_temp_mean, y = site))+
  geom_point(size = 4, color = "navyblue")+
  geom_errorbarh(aes(xmin = wtr_temp_mean-wtr_temp_sd, xmax = wtr_temp_mean + wtr_temp_sd), size = 0.25, height = 0.25)+
  geom_vline(xintercept = 0, linetype = 2)+
  labs(y = NULL, x = "Minimum Water Temperature Effect")+
  theme_classic()+
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 16))

# GDD
ggplot(temp_effects, aes(x = gdd_mean, y = site))+
  geom_point(size = 4, color = "orange")+
  geom_errorbarh(aes(xmin = gdd_mean-gdd_sd, xmax = gdd_mean + gdd_sd), size = 0.25, height = 0.25)+
  geom_vline(xintercept = 0, linetype = 2)+
  labs(y = NULL, x = "Growing Degree Day Effect")+
  theme_classic()+
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 16))


# GDD Quad
ggplot(temp_effects, aes(x = gdd_quad_mean, y = site))+
  geom_point(size = 4, color = "magenta")+
  geom_errorbarh(aes(xmin = gdd_quad_mean-gdd_quad_sd, xmax = gdd_quad_mean + gdd_quad_sd), size = 0.25, height = 0.25)+
  geom_vline(xintercept = 0, linetype = 2)+
  labs(y = NULL, x = "Quadratic Term Growing Degree Day Effect")+
  theme_classic()+
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 16))

```

# Buoy vs. Onset GDD
- compare buoy GDD calculation vs. onset hobo GDD
```{r}
# buoy_GDD <- read_csv("00_Data_files/Covariate_analysis_data/buoy_growing_degree_days.csv")
# onset_GDD <- read_csv("00_Data_files/Covariate_analysis_data/onset_watertemp_all_NB.csv")
# 
# plot(gdd_sum ~ date, data = buoy_GDD)
# points(gdd_sum ~ date, data = onset_GDD, col = "red")
# 
# plot(buoy_GDD$gdd_sum ~ onset_GDD$gdd_sum)
# 
# onset_GDD1 <- onset_GDD %>% 
#   mutate(dayofyr = yday(date), year = year(date)) %>% 
#   select(date, dayofyr, year, gdd_sum)
# 
# buoy_GDD1 <- buoy_GDD %>% 
#   mutate(dayofyr = yday(date), year = year(date)) 
# 
# GDD_join <- full_join(onset_GDD1, buoy_GDD1, by = c("date", "dayofyr", "year")) %>% 
#   rename(onset_gdd = gdd_sum.x, buoy_gdd = gdd_sum.y) %>% 
#   pivot_longer(4:5, names_to = "sensor_type", values_to = "gdd")
# 
# ggplot(data = GDD_join, aes(x = dayofyr, y = gdd, color = sensor_type)) +
#   geom_point() +
#   facet_wrap(~year)
# 
# ggplot(data = GDD_join, aes(x = gdd, y = gdd, color = sensor_type)) +
#   geom_point() +
#   facet_wrap(~year)
```

# Pre 2009 water temp data for prior
```{r}
wtr_pre <- wtr %>% 
  filter(year %in% 2007:2008) %>% 
  filter(site != "HerrickCoveSouth")
  
wtr_hc <- wtr %>% 
  filter(year==2008) %>% 
  filter(site=="HerrickCoveSouth")

unique(wtr_pre$site)
names(wtr_pre)

# Bad HC data pre 	2008-06-18 11:00:00
# Jump @ 2008-06-19 11:00:00, 12, 13

wtr_hc2 <- wtr_hc %>% 
  mutate(temp_degC2 = ifelse(datetime < "2008-06-18 06:00:00", NA, temp_degC)) %>% 
  mutate(temp_degC3 = ifelse(datetime > "2008-06-19 05:00:00" & datetime < "2008-06-19 09:00:00", NA, temp_degC2))



ggplot(wtr_hc2, aes(y = temp_degC3, x = dayofyr, color = site))+
  geom_point()+
  geom_line()+
  facet_wrap(~year)
```


# Load PRISM data all sites
```{r}
# files <- dir("00_Data_files/PRISM_precip_data_all_sites", pattern = "\\.csv$", full.names = TRUE)
# head(files)
# df_list <- vector("list", length(files))
# 
# # Full data
# for (i in seq_along(files)) {
#         df_list[[i]] <- read_csv(files[[i]], skip = 10,
#                    col_types = list(Date = col_date(format = ""),
#                                     `ppt (mm)` = col_double(),
#                                     `tmin (degrees C)` = col_double(),
#                                     `tmean (degrees C)` = col_double(),
#                                     `tmax (degrees C)` = col_double()))
#         
#         colnames(df_list[[i]]) <- c("date", "precip_mm", "air.tempC_min", "air.tempC_mean", "air.tempC_max")
# 
#         df_list[[i]] <-  df_list[[i]] %>% 
#                 select(date, precip_mm) %>%
#                 mutate(site = i) %>% 
#                 mutate(year = year(date)) %>%
#                 mutate(month = month(date)) %>%
#                 filter(year %in% 2009:2016) %>%
#                 filter(month %in% 5:10)
#        
# }
# 
# precip_all_sites <- bind_rows(df_list)
# str(precip_all_sites)
# ```
# 
# # Compare PRISM precip data for all sites
# ```{r}
# #Rename sites
# precip_all_sites1 <- precip_all_sites %>% 
#   mutate(site_name = ifelse(site==1, "NorthSunapeeHarbor", site)) %>% 
#   mutate(site_name = ifelse(site==2, "SouthOfTheFells", site_name)) %>% 
#   mutate(site_name = ifelse(site==3, "HerrickCoveSouth", site_name)) %>% 
#   mutate(site_name = ifelse(site==4, "Newbury", site_name)) %>% 
#   mutate(dayofyr = yday(date))
# ```
# 
# # Plot precip all sites
# ```{r}
# ggplot(data = precip_all_sites1, aes(x = dayofyr, y = precip_mm, color = site_name))+
#   geom_line()+
#   geom_point(data = gloeo_final, aes(x = dayofyr, y = totalperL, color = site), shape = 17, size = 3)+ #Try adding in gloeo data on top
#   scale_color_brewer(palette = "Spectral")+
#   facet_wrap(~year, scales = "free_y")
# ```
# 
# ```{r}
# # Join with sampling dates 
# #read in sampling dates
# sampling_dates <- read_csv("./00_Data_files/sampling_dates_NB.csv")
# 
# # Filter precip data for sampling dates
# precip_all_sites2 <- precip_all_sites1 %>%
#   filter(date %in% sampling_dates$date)
# 
# ggplot(data = precip_all_sites2, aes(x = dayofyr, y = precip_mm, color = site_name))+
#   geom_line()+
#   geom_point()+
#   geom_point(data = gloeo_final, aes(x = dayofyr, y = totalperL, color = site), shape = 17, size = 3)+#Try adding in gloeo data on top
#   scale_color_brewer(palette = "Spectral")+
#   facet_wrap(~year, scales = "free_y")
```

# Wind data wrangling
```{r}
# # Load buoy wind data into R ####
# buoy_wind <- read_csv("./00_Data_files/EDI_data_clones/2007_e2019_wind_L1.csv", col_types = cols(
#   datetime = col_datetime(format = ""),
#   location = col_character(),
#   WindDir_deg = col_double(),
#   WindSp_ms = col_double(),
#   AveWindDir_deg = col_double(),
#   AveWindSp_ms = col_double(),
#   MaxWindSp_ms = col_double(),
#   MaxWindDir_deg = col_double(),
#   wind_flag = col_character()))
# ```
# 
# # Check and filter wind data
# ```{r}
# buoy_wind1 <- buoy_wind %>%
#   mutate(date = as_date(datetime)) %>%
#   mutate(year = year(datetime)) %>%
#   mutate(month = month(datetime)) %>%
#   filter(year %in% 2009:2016) %>%
#   filter(month %in% 5:10)
# 
# print(unique(buoy_wind1$wind_flag)) #NA, so no flags to worry about
# 
# # Use instantaneous data for 2011-2012 - no Avg wind speed data
# buoy_wind_inst <- buoy_wind1 %>%
#   filter(year %in% 2011:2012) %>%
#   select(1,10:12,2:4) %>%
#   rename(AveWindSp_ms = WindSp_ms, AveWindDir_deg = WindDir_deg) # rename to be able to join columns
# 
# # Use average wind data for the rest of the years - no instantaneous data for 2015-2016
# buoy_wind_avg <- buoy_wind1 %>%
#   filter(year %in% c(2009:2010, 2013:2016)) %>%
#   select(1,10:12,2,5:6)
# 
# # Join wind datasets
# buoy_wind2 <- bind_rows(buoy_wind_inst, buoy_wind_avg) %>%
#   select(-c(date, year, month))
# 
# # check data
# plot(AveWindSp_ms ~ datetime, data = buoy_wind2)
# summary(buoy_wind2)
# ```
# 
# # Combine wind data with full time series
# ```{r}
# full_datetime <- seq(from = ymd_hms("2009-05-21 00:00:00"), to = ymd_hms("2016-10-05 23:00:00"), by = dminutes(10))
# full_datetime.df <- as_tibble_col(full_datetime, column_name = "datetime")
# 
# buoy_wind3 <- left_join(full_datetime.df, buoy_wind2, by = "datetime") %>%
#   rename(date = datetime) # open air function needs date column so renamed datetime
# 
# # Calculate hourly mean of wind speed and wind direction data ####
# windsp_hourly_mean <- timeAverage(buoy_wind3, avg.time = "hour", data.thresh = 75, statistic = "mean", start.date = "2009-05-21 00:00:00", end.date = "2016-10-05 23:00:00", interval = "10 min", vector.ws = T)
# ```
# 
# 
# # Add indicator variable for hourly wind direction
# ```{r}
# # If wind is blowing towards Herrick Cove 180-359°, use 1, if blowing away 0-179°, use 0
# # For newbury if wind is blowing towards the southern end of the lake (coming from the north) use 1 for 315-45°
# windsp_hourly_mean_filter <- windsp_hourly_mean %>%
#   mutate(AveWindDir_cove = ifelse(AveWindDir_deg >= 315 | AveWindDir_deg < 46, 1, 0))
# ```
# 
# 
# # Calculate daily summaries from hourly average data
# ```{r}
# 
# windsp_daily_mean <- timeAverage(windsp_hourly_mean_filter, avg.time = "day", data.thresh = 50, statistic = "mean", start.date = "2009-05-21 00:00:00", end.date = "2016-10-05 23:00:00", interval = "60 min", vector.ws = T)
# 
# windsp_daily_median <- timeAverage(windsp_hourly_mean_filter, avg.time = "day", data.thresh = 50, statistic = "median", start.date = "2009-05-21 00:00:00", end.date = "2016-10-05 23:00:00", interval = "60 min", vector.ws = T)
# 
# windsp_daily_min <- timeAverage(windsp_hourly_mean_filter, avg.time = "day", data.thresh = 50, statistic = "min", start.date = "2009-05-21 00:00:00", end.date = "2016-10-05 23:00:00", interval = "60 min", vector.ws = T)
# 
# windsp_daily_max <- timeAverage(windsp_hourly_mean_filter, avg.time = "day", data.thresh = 50, statistic = "max", start.date = "2009-05-21 00:00:00", end.date = "2016-10-05 23:00:00", interval = "60 min", vector.ws = T)
# 
# windsp_daily_sd <- timeAverage(windsp_hourly_mean_filter, avg.time = "day", data.thresh = 50, statistic = "sd", start.date = "2009-05-21 00:00:00", end.date = "2016-10-05 23:00:00", interval = "60 min", vector.ws = T)
# 
# windsp_daily_sum <- timeAverage(windsp_hourly_mean_filter, avg.time = "day", data.thresh = 50, statistic = "sum", start.date = "2009-05-21 00:00:00", end.date = "2016-10-05 23:00:00", interval = "60 min", vector.ws = T)
# 
# #Bind summaries together
# windsp_daily_summary <- bind_cols(windsp_daily_mean, windsp_daily_median[,3], windsp_daily_min[,3], windsp_daily_max[,3], windsp_daily_sd[,3], windsp_daily_sum[,3:4])
# colnames(windsp_daily_summary)[3:10] <- c("AveWindSp_ms_mean","AveWindDir_cove_mean","AveWindSp_ms_median","AveWindSp_ms_min","AveWindSp_ms_max","AveWindSp_ms_sd", "AveWindSp_ms_sum", "AveWindDir_cove_sum")
# ```
# 
# # Wind data visualization
# ```{r}
# # filter for sampling months
# windsp_daily_summary2 <- windsp_daily_summary %>% 
#   mutate(year = year(date)) %>% 
#   mutate(month = month(date)) %>% 
#   filter(month %in% 5:10) %>% 
#   mutate(doy = yday(date))
# 
# ggplot(data = windsp_daily_summary2, aes(x = doy, y = AveWindSp_ms_mean))+
#   geom_line()+
#   geom_point()+
#   facet_wrap(~year)
# 
# sum(is.na(windsp_daily_summary2$AveWindSp_ms_mean))
# ```
# 
# # Wind rose
# ```{r}
# # wind rose function in openair package
# 
# # Daily summary of wind data
# windRose(windsp_daily_summary2, ws = "AveWindSp_ms_mean", wd = "AveWindDir_deg", type = "year",paddle = F , breaks = c(0,2,4,6,8,10))
# 
# windRose(windsp_daily_summary2, ws = "AveWindSp_ms_mean", wd = "AveWindDir_deg", type = "month",paddle = F , breaks = c(0,2,4,6,8,10))
# 
# # Compare with raw 10 min data
# buoy_wind4 <- buoy_wind3 %>% 
#   mutate(year = year(date)) %>% 
#   mutate(month = month(date)) %>% 
#   filter(month %in% 5:10)
# 
# # 10 min raw wind data
# windRose(buoy_wind4, ws = "AveWindSp_ms", wd = "AveWindDir_deg", type = "year",paddle = F , breaks = c(0,2,4,6,8,10))
# 
# windRose(buoy_wind4, ws = "AveWindSp_ms", wd = "AveWindDir_deg", type = "month",paddle = F , breaks = c(0,2,4,6,8,10))
# ```
# 
# Wind rose for high years 2013-Midge, 2010 high for both, 2009 highest for Newbury
# ```{r}
# 
# # use 10 min data but filter for year & month
# buoy_wind5 <- buoy_wind4 %>% 
#   filter(year == 2013) %>% 
#   filter(month == 9) %>% 
#   mutate(day = day(date)) %>% 
#   mutate(day_fac = as.factor(day(date))) # functions needs factor for figure
#   #filter(day > 12)
# 
# str(buoy_wind5)
# 
# windRose(buoy_wind5, ws = "AveWindSp_ms", wd = "AveWindDir_deg", type = "day_fac",paddle = F, breaks = c(0,2,4,6,8,10))


```

# Sunapee Master data LMP
```{r}
sunapee_master_http <- 'https://raw.githubusercontent.com/Lake-Sunapee-Protective-Association/LMP/main/master%20files/LSPALMP_1986-2020_v2021-03-29.csv'

sunapee_master <- read_csv(sunapee_master_http,
                           col_types = cols(.default = col_character()))
unique(sunapee_master$station) #Midge (site110) and Newbury (site90)    
unique(sunapee_master$site_type)

#subset for sites 110 and 90, secchi and TP, 2009-2016, May-Oct
sunapee_filter <- sunapee_master %>%
  filter(station == 110|station == 90) %>%
  filter(parameter == 'secchidepth_m' | parameter == 'TP_mgl') %>%
  filter(!is.na(value)) %>%
  mutate(date = as.Date(date),
         month = as.numeric(format(date, '%m')),
         value = as.numeric(value)) %>%
  filter(date >= as.Date('2009-01-01') & date < as.Date('2017-01-01') & month >= 5 & month < 11)

# convert TP to ugl
sunapee_filter <- sunapee_filter %>%
  mutate(value = case_when(parameter == 'TP_mgl' ~ value * 1000,
                           TRUE ~ value),
         parameter = case_when(parameter == 'TP_mgl' ~ 'TP_ugl',
                               TRUE ~ parameter))

```

# Plot TP data over time - y-scale varies
Y-scale set to free for each year
```{r}
names(sunapee_filter)
sunapee_TP <- sunapee_filter %>% 
  filter(parameter == "TP_ugl")

ggplot(data = sunapee_TP, aes(x = date, y = value, shape = station, color = depth_m))+
  geom_point()+
  geom_line()
```

# Table with Hindcasting results for week 1
```{r}
# hindcast_output_analysis_wk_1 <- read_csv("6_Output_analysis/hindcast_output_analysis_wk_1.csv")
# names(hindcast_output_analysis_wk_1)
# 
# hindcast_wk1 <- hindcast_output_analysis_wk_1 %>% 
#   mutate(changePL = pred_loss-min(pred_loss)) %>% 
#   select(model_name, RMSE, pred_SD, pred_loss, changePL, coverage, peak_timing, mean_bias)
# 
# header <- c("Model Name", "RMSE", "Predictive S.D.", "Predictive Loss","Change PL",  "% Coverage", "Peak Timing", "Bias")
# 
# kable(hindcast_wk1, digits = 2, col.names = header, caption = "Table 1. Newbury 1-Week Hindcast", align = "lcccccc")
# ```
# 
# # Table with Hindcasting results for week 4
# ```{r}
# hindcast_output_analysis_wk_4 <- read_csv("6_Output_analysis/hindcast_output_analysis_wk_4.csv")
# names(hindcast_output_analysis_wk_4)
# 
# hindcast_wk4 <- hindcast_output_analysis_wk_4 %>% 
#   mutate(changePL = pred_loss-min(pred_loss)) %>% 
#   select(model_name, RMSE, pred_SD, pred_loss, changePL, coverage, peak_timing, mean_bias)
# 
# header <- c("Model Name", "RMSE", "Predictive S.D.", "Predictive Loss","Change PL",  "% Coverage", "Peak Timing", "Bias")
# 
# kable(hindcast_wk4, digits = 2,  col.names = header, caption = "Table 2. Newbury 4-Week Hindcast", align = "lccccccc")
```

